---
title: "Transitions analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Raw data of counts of each cluster at each time point

```{r}
nf=8
nams=c("E1",	"E2",	"E3",	"pEMT1",	"pEMT2",	"pEMT3",	"M","pMET")

#  Data Set 1, with rowsum normalized to 10,0000
r0 = c(4417, 2240, 2080, 645, 137, 333, 105, 43)
r1 = c(158, 1730, 1145, 2932, 939, 2268, 670, 157)
r2 = c(8, 162, 128, 842, 1835, 2092, 4653, 280)
r3 = c(2, 32, 12, 283, 909, 1147, 7285, 330)
r4 = c(27, 111, 9, 838, 1532, 165, 3826, 3492)
r5 = c(216, 1107, 28, 1365, 1043, 301, 4156, 1785)
r6 = c(1085, 2470, 53, 1027, 783, 198, 2240, 2144)


rcount1 = t(rbind(r0,r1,r2,r3,r4,r5,r6))

dimnames(rcount1)=list(nams,c("0","2","6","10","w2","w6","w10"))

rprop1=scale(rcount1,F,colSums(rcount1))  #make colsums equal to 1

```





```{r results=FALSE, warning=FALSE, message=FALSE}
library("tidyverse")
library("wesanderson")
```



# Load in results from bootstrap simulations

```{r}
boot_files <- file.path("boot_samples", list.files(path="boot_samples", pattern="boot_res"))
boot_list <- lapply(boot_files, readRDS)
```

```{r}
lambdas_1se <- sapply(boot_list, function(x) x$lambda_1se)
```

```{r}
nnzeros_p1 <- sapply(boot_list, function(x) sum(x$train_res$post_res$transition_P[[1]] >=10^(-4)))
nnzeros_p2 <- sapply(boot_list, function(x) sum(x$train_res$post_res$transition_P[[2]] >=10^(-4)))

lambda_cv_nnzero_df <- rbind(data.frame(lambda=lambdas_1se, nnzeros = nnzeros_p1, transition="P1"), 
                             data.frame(lambda=lambdas_1se, nnzeros = nnzeros_p2, transition="P2"))
```


# Plot of number of nonzeros for the two estimated transition matrices versus chosen regularization parameter (each point represents one bootstrap replication)

```{r fi.width=10}
ggplot(lambda_cv_nnzero_df, aes(x=lambda, y=nnzeros)) + geom_jitter() + theme_bw() + facet_grid(.~transition)
```



```{r}
df_boot <- data.frame()
K_clusters <- 8
for (i in 1:K_clusters){
  for (j in 1:K_clusters){
    #if (diff_P1_P2_sparsity[i,j] == 1){
      p1_samples <- sapply(boot_list, function(u) u$train_res$post_res$transition_P[[1]][i,j])
      p2_samples <- sapply(boot_list, function(u) u$train_res$post_res$transition_P[[2]][i,j])
      df_boot <- rbind(df_boot,
                       data.frame(P1=p1_samples,
                                  P2=p2_samples,
                                  to=nams[i], 
                                  from=nams[j], 
                                  boot_id = 1:length(boot_list)))
  }
}
    
    
```


```{r}
df_boot_m <- gather(df_boot, transition, probability, -c(to,from,boot_id)) 

stage_cluster_levels <-  unlist(lapply(unique(df_boot_m$to), function(x) paste0(x, c(" (P1)"," (P2)"))))

df_boot_m <- mutate(df_boot_m, to_stage = paste0(to, " (",transition,")")) %>% 
             mutate(to_stage=factor(to_stage,levels=stage_cluster_levels)) %>%
             mutate(from = factor(from, levels=levels(to))) 

df_boot_m_from <- mutate(df_boot_m, from= factor(paste0( "from: ", from), levels=paste0("from: ", levels(from))))
df_boot_m_to <-  mutate(df_boot_m, to= factor(paste0( "to: ", to), levels=paste0("to: ", levels(to))))
```

```{r}
# sanity check:
group_by(df_boot_m_from, from, boot_id, transition) %>%
             summarize(sum_prob = sum(probability)) %>% 
             pull(sum_prob) %>%
             range
```

# Bootstrap results

```{r fig.width=12,fig.height=7}
p_boot <- ggplot(df_boot_m_from, aes(x=to, y=probability,fill=transition)) + geom_boxplot() + 
  facet_wrap(~from,ncol=2, scales="free_y") + theme_bw(base_size=14) + 
  scale_fill_manual(values=wes_palette(n=2,name="Chevalier1"))+
  theme(axis.text.x=element_text(angle=90, hjust=1))

p_boot
```

```{r eval=FALSE}
ggsave("bootstrap_plot.pdf", p_boot, width=12, height=7)
```

## Now also transpose, i.e. show to->from

```{r fig.width=12,fig.height=7}
ggplot(df_boot_m_to, aes(x=from, y=probability,fill=transition)) + geom_boxplot() + 
  facet_wrap(~to,ncol=2, scales="free_y") + theme_bw(base_size=14) + 
  scale_fill_manual(values=wes_palette(n=2,name="Chevalier1"))+
  theme(axis.text.x=element_text(angle=90, hjust=1))
```

## Now generate graphs


```{r results=FALSE, warning=FALSE, message=FALSE}
library(igraph)
library(ggraph)
set_graph_style(plot_margin = margin(3,3,3,3))
```

```{r}
library(cowplot)
```

```{r}
# function that plots the graph
plot_adjacency <- function(mat, cutoff=10^{-3}){

  Net1 <- graph_from_adjacency_matrix(t(mat*(mat  > cutoff)),
                                      mode="directed",
                                      weighted=TRUE)

  E(Net1)$pretty_weight <- as.character(round(edge_attr(Net1)$weight,3))
  for (i in 1:length(E(Net1))){
    E_tmp = E(Net1)[i]
    if (ends(Net1,E_tmp)[1] ==  ends(Net1,E_tmp)[2]){
      E(Net1)$pretty_weight[i] = ""
    }
  }


  cluster_layout <- as.data.frame( rbind( c(1,2),
                                          c(0.75,1),
                                          c(1,0),
                                          c(2.25,2.5),
                                          c(2,1),
                                          c(2.25,-0.5),
                                          c(3,1.75),
                                          c(3.5,0.5)))
  colnames(cluster_layout) <- c("x","y")



  ggraph(Net1, layout= cluster_layout) +
    geom_edge_fan(aes(end_cap = circle(5.5, 'mm'), label=pretty_weight),
                  angle_calc = 'along',
                  label_dodge = unit(2.5, 'mm'),
                  arrow =  arrow(length = unit(6, "pt"), type = "closed")) +
    geom_node_label(aes(label = name))
}
```

# Find medoids and plot graph
```{r}
n_boot <- length(boot_files)
dist_mat <- matrix(NA, n_boot, n_boot)
for (i in 1:(n_boot-1)){
  for (j in (i+1):n_boot){
    P_i_1 <- boot_list[[i]]$train_res$post_res$transition_P[[1]]
    P_i_2 <- boot_list[[i]]$train_res$post_res$transition_P[[2]]
    P_j_1 <- boot_list[[j]]$train_res$post_res$transition_P[[1]]
    P_j_2 <- boot_list[[j]]$train_res$post_res$transition_P[[2]]
    dist_mat[i,j] <- sum( abs(P_i_1 - P_j_1) + abs(P_i_2 - P_j_2))
    dist_mat[j,i] <- dist_mat[i,j]
  }
}

score_vec <- sapply( 1:n_boot, function(i) sum(dist_mat[i,], na.rm=TRUE))
idx_medoid <- which.min(score_vec)
P1_medoid_L1 <- boot_list[[idx_medoid]]$train_res$post_res$transition_P[[1]]
P2_medoid_L1 <- boot_list[[idx_medoid]]$train_res$post_res$transition_P[[2]]
```




```{r fig.width=17, fig.height=7}
plot_grid( plot_adjacency(P1_medoid_L1, cutoff=0.01), 
           plot_adjacency(P2_medoid_L1, cutoff=0.01),
           nrow=1)
```

# Session information

```{r}
sessionInfo()
```
